---
title: "KPCA Demonstration on Synthetic Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{kpca_synthesis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 5, fig.width = 7, fig.align = "center"
)
```

```{r setup}
library(dtrans)
```

Two-dimentional data is generated of being non-linearly separated.

```{r gen_synthesis_data, echo=FALSE}
set.seed(1L)
n <- 100L
theta <- runif(n, 0, 2 * pi)
r <- rlnorm(n, meanlog = log(6), sdlog = 0.1)
df1 <- data.frame(x=r * cos(theta), y=r * sin(theta))

set.seed(1L)
n <- 100L
theta <- runif(n, 0, 2 * pi)
r <- rlnorm(n, meanlog = log(2), sdlog = 0.1)
# plot(r * cos(theta), r * sin(theta), asp = 1)
df2 <- data.frame(x=r * cos(theta), y=r * sin(theta))
df <- rbind(df1,df2)
df$y <- df$y+abs(min(df$y))
df$x <- df$x+abs(min(df$x))

df <- df[sample(1:nrow(df)),]
rownames(df) <- 1:nrow(df)

plot(df)
```

Three *transformer* objects are created to reduce 2D data into 1D by PCA, NMF, and KPCA. There are no clear separation with PCA and NMF, while KPCA provides a clear distinction of two possible groups. So, we will groups point based on the condition on Component 1 value of greater versus smaller than zero.

```{r, fig.height = 3}
pca <- transformer.pca(df, components = 1)
nmf <- transformer.nmf(df, components = 1)
kpca <- transformer.kpca(df, components = 1)

par(mfrow=c(1, 3))
plot(pca, main="PCA")
plot(nmf, main="NMF")
plot(kpca, main="KPCA")
```

Then, we plot the 3 *transformer* object above and color the by the groups created.

```{r, fig.height = 3}
grp1_idx <- which(kpca$x[,'PC1'] > 0)
grp2_idx <- which(kpca$x[,'PC1'] < 0)

df['color'] <- "black"
df[grp2_idx, 'color'] <- "red"
par(mfrow=c(1, 3))
plot(pca, col=df$color, main="PCA")
plot(nmf, col=df$color, main="NMF")
plot(kpca, col=df$color, main="KPCA")

```

The coloring by groups is assigned to original data. The two groups shows a good separation by two donut-shape clusters.

```{r}
plot(df[,1:2], col=df$color)
```
