---
title: "my-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dtrans)
```

This package is about transformer object which is created by different algorithms including PCA, NMF, and Kernel PCA. It transforms the high 
dimensionality dataset into a lower one to serve the visualization purpose or dimension reduction tasks.

Firstly, the object will be simply created with the given dataset:

```{r}
data("iris")
#pca <- transformer.pca(iris[,1:4])
#nmf <- transformer.nmf(iris[,1:4])
#kpca <- transformer.kpca(iris[,1:4])

```

The created objects are class 'Transformer' with attributes:

```{r}
#attributes(pca)
```

* x: transformed data
* sdev: 
* coef: coefficient of original features to each fitted component
* center: value used as center to scale data, default is mean of each original feature
* scale: value used to scale the data before transforming, default is std of original feature
* technique: indicate the algorithm used to create the object
* data: original data

